Estimating Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks

### 2024-07-09 paper/coverletter.tex
Update final version

We submit our manuscript ``Estimating Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks'' for publication in Archives of Acoustics. Authored by Paweł Antoniuk (corresponding author) and Sławomir Krzysztof Zieliński, this paper presents a method for estimating spatial parameters in binaural audio recordings using a novel ensemble-based approach employing a spectrogram-based deep learning network technique.

\begin{itemize}
  \item We submit this paper to Archives of Acoustics as its content aligns with the journal's scope, particularly in \textbf{speech processing and communication systems}.
  \item We confirm this manuscript is original, unpublished, and not under consideration elsewhere.
  \item We recommend the following reviewers:
        \begin{itemize}
          \item \textbf{prof. dr hab. inż. Bożena Kostek} \\*
          Wydział Elektroniki, Telekomunikacji i Informatyki \\*
          Gdańsk University of Technology \\*
          ul. Gabriela Narutowicza 11/12 \\*
          80-233 Gdańsk \\*
          Poland \\*
          Email: \href{mailto:bozenka@sound.eti.pg.gda.pl}{bozenka@sound.eti.pg.gda.pl}
          \item \textbf{prof. dr hab. inż. Piotr Kleczkowski} \\*
          Wydział Inżynierii Mechanicznej i Robotyki \\*
          AGH University of Krakow \\*
          al. Mickiewicza 30 \\*
          30-059 Kraków \\*
          Poland \\*
          Email: \href{mailto:kleczkow@agh.edu.pl}{kleczkow@agh.edu.pl}
          \item \textbf{dr hab. inż. Grzegorz Szwoch} \\*
          Wydział Elektroniki, Telekomunikacji i Informatyki \\*
          Gdańsk University of Technology \\*
          ul. Gabriela Narutowicza 11/12 \\*
          80-233 Gdańsk \\*
          Poland \\*
          Email: \href{mailto:grzegorz.szwoch@pg.edu.pl}{grzegorz.szwoch@pg.edu.pl}
        \end{itemize}
  \item We propose classifying this manuscript as a \textbf{``research paper''} due to its empirical content.
  \item While the main content is within 17 pages, the manuscript exceeds the recommended 20-page limit to provide comprehensive context, detailed methodology, and extensive references. We believe this additional content is crucial for a full understanding of our research and to allow others to replicate our work.
\end{itemize}

>>>>

We submit our manuscript ``Estimating Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks'' for publication in Archives of Acoustics. Authored by Paweł Antoniuk (corresponding author) and Sławomir Krzysztof Zieliński, this paper presents a method for estimating spatial parameters in binaural audio recordings using a novel ensemble-oriented approach employing a spectrogram-based deep learning network technique.

\begin{itemize}
  \item We submit this paper to Archives of Acoustics as its content aligns with the journal's scope, particularly in \textbf{speech processing and communication systems}.
  \item We confirm this manuscript is original, unpublished, and not under consideration elsewhere.
  \item We recommend the following reviewers:
        \begin{itemize}
          \item \textbf{prof. dr hab. inż. Bożena Kostek} \\*
          Faculty of Electronics, Telecommunications and Informatics \\*
          Gdańsk University of Technology \\*
          ul. Gabriela Narutowicza 11/12 \\*
          80-233 Gdańsk \\*
          Poland \\*
          Email: \href{mailto:bozenka@sound.eti.pg.gda.pl}{bozenka@sound.eti.pg.gda.pl}
          \item \textbf{prof. dr hab. inż. Piotr Kleczkowski} \\*
          Faculty of Mechanical Engineering and Robotics \\*
          AGH University of Krakow \\*
          al. Mickiewicza 30 \\*
          30-059 Kraków \\*
          Poland \\*
          Email: \href{mailto:kleczkow@agh.edu.pl}{kleczkow@agh.edu.pl}
          \item \textbf{dr hab. inż. Grzegorz Szwoch} \\*
          Faculty of Electronics, Telecommunications and Informatics \\*
          Gdańsk University of Technology \\*
          ul. Gabriela Narutowicza 11/12 \\*
          80-233 Gdańsk \\*
          Poland \\*
          Email: \href{mailto:grzegorz.szwoch@pg.edu.pl}{grzegorz.szwoch@pg.edu.pl}
          \item \textbf{Professor Philip J. B. Jackson} \\*
          University of Surrey \\*
          Guildford GU2 7XH \\*
          United Kingdom \\*
          Email: \href{mailto:p.jackson@surrey.ac.uk}{p.jackson@surrey.ac.uk}
        \end{itemize}
  \item We propose classifying this manuscript as a \textbf{``research paper''} due to its empirical content.
  \item While the main content is within 17 pages, the manuscript exceeds the recommended 20-page limit to provide comprehensive context, detailed methodology, and extensive references. We believe this additional content is crucial for a full understanding of our research and to allow others to replicate our work.
\end{itemize}

### 2024-07-09 paper/main.tex
Update final version

\doublespacing
\defbibenvironment{bibliography}
  {\begin{enumerate}}
  {\end{enumerate}}
  {\item}
\AtEveryBibitem{%
  \clearfield{note}%
}
\renewcommand\Affilfont{\itshape\centering}
\renewenvironment{abstract}{}{}
\newcommand{\keywords}[1]{
  \noindent\begin{center}\begin{minipage}{\textwidth}
  \small\textbf{Keywords:} #1
  \end{minipage}\end{center}
}
\DefineBibliographyStrings{english}{%
  andothers = {\textit{et al\adddot}},
}
\DeclareFieldFormat{doi}{%
  \href{https://doi.org/#1}{#1}%
}
\renewbibmacro*{doi+eprint+url}{%
  \iftoggle{bbx:doi}
    {\printfield{doi}}
    {}%
  \newunit
  \iftoggle{bbx:eprint}
    {\usebibmacro{eprint}}
    {}%
  \newunit
  \iftoggle{bbx:url}
    {\usebibmacro{url+urldate}}
    {}}

@@

\title{Estimating Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks}
\author[1,*]{Paweł Antoniuk\orcidlink{0000-0002-0914-8920}}
\author[1]{Sławomir Zieliński\orcidlink{0000-0002-3205-974X}}
\affil[1]{Faculty of Computer Science, Białystok University of Technology \protect\\ Wiejska 45A, 15-351 Białystok, Poland}
\affil[*]{Corresponding Author Email: pawel.antoniuk@sd.pb.edu.pl}
\date{}

@@

\begin{abstract}
  Binaural audio technology has been in existence for many years. However, its popularity has significantly increased over the past decade as a consequence of advancements in virtual reality and streaming techniques. Along with its growing popularity, the quantity of publicly accessible binaural audio recordings has also expanded. Consequently, there is now a need for automated and objective measurements of spatial content information, with ensemble location and width being the most prominent. This study presents a novel method for estimating these ensemble parameters in binaural recordings of music. For this purpose, a dataset of 23,040 binaural recordings was synthesized from 192 publicly-available music recordings using 30 head-related transfer functions. The synthesized excerpts were then used to train a multi-task spectrogram-based convolutional neural network model, aiming to estimate the ensemble location and width for unseen recordings. The results indicate that a model for estimating ensemble parameters can be successfully constructed with low prediction errors: $4.76\degree$ ($\pm0.10\degree$) for ensemble location and $8.57\degree$ ($\pm0.19\degree$) for ensemble width. The method developed in this study outperforms previous spatiogram-based techniques recently published in the literature and shows promise for future development as part of a novel tool for binaural audio recordings analysis.
\end{abstract}

@@

Humans tend to localize groups of sound sources rather than individual ones \parencite{bregman_auditory_1990, rumsey_spatial_2002}. Inspired by this fact, the objective of the proposed model is to estimate the location and width of these groups, termed ensembles, instead of the positions of individual sources. This study is unique as it not only developed the method but also tested it on a vast, realistic music corpus. The corpus comprised 23,040 binaural excerpts synthesized using 192 multi-track music recordings (from a repository provided by \textcite{senior_mixing_2023}) and 30 publicly available head-related transfer functions (HRTFs) from various sources (see Table \ref{table:hrtfs} in \nameref{appendix:a_hrtf} for a detailed list). The music recordings covered many different genres, including rock, jazz, pop, and classical music.

@@

The developed method has the potential to be highly beneficial in automated assessment tasks, where a significant number of binaural recordings must be measured or labeled in terms of their spatial content information. This could be utilized in the development of a hypothetical autonomous ``web-crawler bot'' that will collect binaural recordings from publicly accessible repositories and label them according to the spatial properties of the sound sources, such as the location of the music ensemble or the sparsity of audio source positions. This method may also assist audio engineers in objectively assessing and segregating binaural audio recordings with regard to their spatial content.

@@

The number of tracks in multi-track recording ranged from 5 to 62, with median of 9. For each pair of a multi-track recording and an HRTF database, four binaural recordings were synthesized with different random ensemble parameters, namely location $\theta$ and width $\omega$, as defined in Section \ref{subsec:ensemble_definition}. Both parameters were drawn from a uniform random distribution. Furthermore, the tracks of the input multi-track recordings were randomly assigned to sound source positions ($\theta_i$) to enhance the diversity of the final binaural corpora. Before the synthesis, the signals in each track were equalized to $-23$ LUFS, in accordance with the ITU-R BS.1770-5 (\citeyear{noauthor_itu-r_2023}) recommendation.

@@

Prior being input into the model, the binaural recordings of music were transformed into magnitude spectrograms. Although spectrograms do not directly provide information that can be translated into ensemble features, especially ensemble width, the goal of this task was to reduce the number of independent variables compared to the raw audio signal by extracting more compressed and informative data in the frequency domain. This step was necessary to decrease the likelihood of overfitting, reduce the number of examples needed to train the model, and thereby lower the overall computational power requirements. It is worth mentioning, however, that recent methods have shown that CNNs are suitable for end-to-end audio localization without the spectrogram extraction step, as shown by \textcite{vera-diaz_towards_2018, vecchiotti_end--end_2019}.

@@

The topology described in the previous section resulted in a model with 216,562 learning parameters. The model training procedure was repeated 10 times, employing the Monte Carlo cross-validation method, as described by \textcite{kuhn_applied_2013}. For each repetition, the entire dataset was randomly divided into two parts: a development set containing two-thirds of the dataset (15,360 recordings) for model construction, and a test set consisting of the remaining one-third of the dataset (7,680 recordings) for its evaluation. This repetition procedure was employed to ensure more reliable and generalizable results by assessing the model's performance across different subsets of the data. Additionally, it helps to account for the inherent variability in neural network training, where slight changes in initial conditions or optimization paths can lead to different model outcomes. While a large and diverse dataset could mitigate this issue, the binaural excerpts used in this study were generated from only 196 multi-track music recordings. This limited source material raised concerns about potential significant variations between the development and test sets in each repetition. These concerns were validated, as the maximum observed difference in MAE between repetitions reached up to $0.85\degree$ for ensemble width.

To ensure that the evaluation process was unbiased, this division was done in such a way that no original multi-track recordings used for synthesis were included in both the development and test sets simultaneously. However, this rule was not applied to HRTFs databases, allowing for the possibility of HRTF information leaking between the development and test sets. This could be seen as a significant limitation of the study. However, it is known that a human auditory system uses a single HRTF represented by ears, head, and torso, only slightly changing throughout the entire life, mainly during infancy \parencite{clifton_growth_1988, king_how_2001}. Therefore, this limitation could be considered in pair how the human auditory system behaves in real life. Nevertheless, it is worth noting that some studies implement HRTF-independent testing for binaural localization models, as demonstrated by \textcite{zielinski_spatial_2022, zielinski_automatic_2022,antoniuk_blind_2023}.

@@

The optimizer was configured with the following hyperparameters: an initial learning rate of $10^{-3}$, a decay rate of $10^{-6}$, and momentum parameters $\beta_1 = 0.9$ and $\beta_2 = 0.999$. Training was conducted using a batch size of 8. An early stopping technique was implemented to prevent overfitting, terminating the process if no improvement was observed on the validation set for 20 consecutive epochs. Consequently, the maximum number of epochs was never reached; instead, training concluded after 25 to 36 epochs, with a median of 27.5 epochs. During the training process, the losses for both outputs were combined additively, ensuring equal weighting of both ensemble features.

The computational work for this study was conducted on a workstation equipped with an RTX Nvidia GeForce 4090 GPU and a 48-core AMD Ryzen ThreadRipper processor. On the software side, MATLAB \parencite{MATLAB} with the Audio Toolbox \parencite{MATLAB_Audio_Toolbox} was used for the binaural recording synthesis, while Python \parencite{python} with the SciPy package \parencite{2020SciPy-NMeth} was used for feature extraction and Keras \parencite{chollet_keras_2015} for training the CNN model. The complete source code for all stages is publicly available on the GitHub repository \parencite{antoniuk_software_2024}. The spectrogram calculation phase required 21 minutes, data partitioning took 34 minutes, and the total training time for all iterations amounted to 40 minutes, making the entire training and evaluation process 95 minutes long.

@@

Figure \ref{fig:actual_vs_predicted_width} compares the actual and predicted ensemble widths for each sample, showing a heteroscedastic relationship between them, with a slight bias towards predicting lower ensemble width values for higher actual widths. This relationship exhibits a strong positive correlation, with a Pearson r of 0.9. Additionally, the results indicate that the model provides more precise predictions for narrower ensemble widths, with an average Mean Absolute Error (MAE) of $5.65^\circ$ for $\omega < 30^\circ$. However, performance deteriorates as the ensemble width increases, resulting in an MAE of $12.44^\circ$ for $\omega > 80^\circ$. This effect is more visible in Figure \ref{fig:mae_width}, which highlights the impact of the actual ensemble width on the precision of prediction. The correlation between the actual ensemble width and prediction error shows a weak positive relationship, with a Pearson r of 0.27

@@

The correlation between actual and predicted ensemble location exhibits a high Pearson r of 0.97, as illustrated in Figure \ref{fig:actual_vs_predicted_location}. In contrast to the ensemble width, no significant relationship is observed between actual location and its prediction error. This finding suggests that the model's ability to localize the center of the ensemble is robust, unaffected by the actual spatial positioning of the ensemble, including lateral locations. Figure \ref{fig:mae_location} corroborates this observation, demonstrating relatively consistent location error across most positions, with minor increases at extreme locations. The negligible correlation (Pearson r = -0.03) between absolute location value and prediction error further supports the model's spatial invariance in performance.

@@

Figure \ref{fig:map_mae_width} reveals a characteristic depression in $\omega \in [30\degree, 60\degree]$ previously shown from a different perspective in Figure \ref{fig:mae_width}. This heatmap highlights another interesting phenomenon in its upper corners as the error in these areas is considerably higher. This indicates that the model's performance for estimating ensemble width is substantially worse at extreme widths and locations, i.e., when both the width and locations are near their maximum investigated values ($|\theta| \approx 45\degree, \omega \approx 90\degree$).

@@

In terms of ensemble location prediction, the novelty of the proposed method makes direct comparison with existing literature challenging. However, its efficacy can be only evaluated indirectly against state-of-the-art individual-source localization techniques. The ensemble location prediction precision (MAE $=4.76\degree$) of the proposed method can be contextualized with the leading-edge binaural localization DeepEar model introduced by \textcite{yang_deepear_2022}. This model reported MAEs of $7.4\degree$ and $2.3\degree$ for multi-source and single-source Angle of Arrival (AoA) estimation, respectively. As another promising example, the WaveLoc-CONV model developed by \textcite{vecchiotti_end--end_2019} demonstrated errors of $0\degree$ in anechoic conditions and $1.7\degree-2.4\degree$ in multi-condition scenarios. However, these results are limited to single-source speech localization, a substantially less complex task than the ensemble location prediction addressed by the proposed method. These experiments, while differing in objectives and datasets, provide valuable context for the proposed method's performance within current DoA and AoA estimation research.

@@

These limitations, however, present opportunities for future research. Despite the current constraints, this study introduces a novel method for describing acoustic scenes in binaural recordings of music, demonstrating substantial potential for advancing binaural audio analysis. The method offers promising prospects for developing innovative tools that can objectively measure and analyze large repositories of binaural audio, focusing on spatial content.

% \clearpage
\section*{Acknowledgments}
The work was supported by the grants from Białystok University of Technology (WI/WI-IiT/3/2022 and WZ/WI-IIT/5/2023) and funded with resources for research by the Ministry of Science and Higher Education in Poland.

>>>>

\doublespacing
\defbibenvironment{bibliography}
  {\begin{enumerate}}
  {\end{enumerate}}
  {\item}
\AtEveryBibitem{%
  \clearfield{note}%
}
\renewcommand\Affilfont{\itshape\centering}
\renewenvironment{abstract}{}{}
\newcommand{\keywords}[1]{
  \noindent\begin{center}\begin{minipage}{\textwidth}
  \small\textbf{Keywords:} #1
  \end{minipage}\end{center}
}
\DefineBibliographyStrings{english}{%
  andothers = {\textit{et al\adddot}},
}
\DeclareFieldFormat{doi}{%
  doi: \href{https://doi.org/#1}{\nolinkurl{#1}}%
}
\renewbibmacro*{doi+eprint+url}{%
  \iftoggle{bbx:doi}
    {\printfield{doi}}
    {}%
  \newunit
  \iftoggle{bbx:eprint}
    {\usebibmacro{eprint}}
    {}%
  \newunit
  \iftoggle{bbx:url}
    {\usebibmacro{url+urldate}}
    {}}

@@

\title{Estimating Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks}
\author[1,*]{Paweł Antoniuk\orcidlink{0000-0002-0914-8920}}
\author[1]{Sławomir K. Zieliński\orcidlink{0000-0002-3205-974X}}
\affil[1]{Faculty of Computer Science, Białystok University of Technology \protect\\ Wiejska 45A, 15-351 Białystok, Poland}
\affil[*]{Corresponding Author Email: pawel.antoniuk@sd.pb.edu.pl}
\date{}

@@

\begin{abstract}
  Binaural audio technology has been in existence for many years. However, its popularity has significantly increased over the past decade as a consequence of advancements in virtual reality and streaming techniques. Along with its growing popularity, the quantity of publicly accessible binaural audio recordings has also expanded. Consequently, there is now a need for automated and objective retrieval of spatial content information, with ensemble location and width being the most prominent. This study presents a novel method for estimating these ensemble parameters in binaural recordings of music. For this purpose, a dataset of 23,040 binaural recordings was synthesized from 192 publicly-available music recordings using 30 head-related transfer functions. The synthesized excerpts were then used to train a multi-task spectrogram-based convolutional neural network model, aiming to estimate the ensemble location and width for unseen recordings. The results indicate that a model for estimating ensemble parameters can be successfully constructed with low prediction errors: $4.76\degree$ ($\pm0.10\degree$) for ensemble location and $8.57\degree$ ($\pm0.19\degree$) for ensemble width. The method developed in this study outperforms previous spatiogram-based techniques recently published in the literature and shows promise for future development as part of a novel tool for binaural audio recordings analysis.
\end{abstract}

@@

Humans tend to localize groups of sound sources rather than individual ones \parencite{bregman_auditory_1990, rumsey_spatial_2002}. Inspired by this fact, the objective of the proposed model is to estimate the location and width of these groups, termed ``ensembles,'' instead of the positions of individual sources. This study is unique as it not only developed the method but also tested it on a relatively large, realistic music corpus. The corpus comprised 23,040 binaural excerpts synthesized using 192 multi-track music recordings (from a repository provided by \textcite{senior_mixing_2023}) and 30 sets of publicly available head-related transfer functions (HRTFs) acquired from various sources (see Table \ref{table:hrtfs} in \nameref{appendix:a_hrtf} for a detailed list). The music recordings covered many different genres, including rock, jazz, pop, and classical music.

@@

The developed method has the potential to be highly beneficial in automated information retrieval tasks, where a significant number of binaural recordings must be analyzed or labeled in terms of their spatial content information. This could be utilized in the development of a hypothetical autonomous ``web-crawler bot'' that will collect binaural recordings from publicly accessible repositories and label them according to the spatial properties of the sound sources, such as the location of the music ensemble or the sparsity of audio source positions. This method may also assist audio engineers in objectively assessing and segregating binaural audio recordings with regard to their spatial content.

@@

The number of tracks in multi-track recording ranged from 5 to 62, with median of 9. For each pair of a multi-track recording and an HRTF database, four binaural recordings were synthesized with different random ensemble parameters, namely location $\theta$ and width $\omega$, as defined in Section \ref{subsec:ensemble_definition}. Both parameters were drawn from a uniform random distribution. Furthermore, the tracks of the input multi-track recordings were randomly assigned to sound source positions ($\theta_i$) to enhance the diversity of the final binaural corpora. Before the synthesis, the signals in each track were equalized to $-23$ LKFS, in accordance with the ITU-R BS.1770-5 (\citeyear{noauthor_itu-r_2023}) recommendation.

@@

Prior being input into the model, the binaural recordings of music were transformed into magnitude spectrograms. Although spectrograms do not directly provide information that can be translated into ensemble features, especially ensemble width, the goal of this task was to reduce the number of independent variables compared to the raw audio signal by extracting more compressed and informative data in the frequency domain. This step was also necessary to decrease the likelihood of overfitting, reducing the number of examples needed to train the model, and thereby lower the overall computational power requirements. It is worth mentioning, however, that recently published studies have shown that CNNs are suitable for end-to-end audio localization without the spectrogram extraction step, as demonstrated by \textcite{vera-diaz_towards_2018, vecchiotti_end--end_2019}.

@@

The topology described in the previous section resulted in a model with 216,562 learning parameters. The model training procedure was repeated 10 times, employing the Monte Carlo cross-validation method, as described by \textcite{kuhn_applied_2013}. For each repetition, the entire dataset was randomly divided into two parts: a development set containing two-thirds of the dataset (15,360 recordings) for model construction, and a test set consisting of the remaining one-third of the dataset (7,680 recordings) for its evaluation. This repetition procedure was employed to ensure more reliable and generalizable results by assessing the model's performance across different subsets of the data. Additionally, it helped to account for the inherent variability in neural network training, where slight changes in initial conditions or optimization paths can lead to different model outcomes. While a large and diverse dataset could mitigate this issue, the binaural excerpts used in this study were generated from only 196 multi-track music recordings. This limited source material raised concerns by these authors about potential significant variations between the development and test sets in each repetition. In hindsight, these concerns were valid, as the maximum observed difference in MAE between repetitions reached up to $0.85\degree$ for ensemble width.

To ensure that the evaluation process was unbiased, the data split was done in such a way that no original multi-track recordings used for synthesis were included in both the development and test sets simultaneously. However, this rule was not applied to HRTFs databases, allowing for the possibility of HRTF information leaking between the development and test sets. This could be seen as a significant limitation of the study. However, it is known that a human auditory system uses a single HRTF represented by ears, head, and torso, only slightly changing throughout the entire life, mainly during infancy \parencite{clifton_growth_1988, king_how_2001}. Therefore, this limitation could be considered in pair how the human auditory system behaves in real life. Nevertheless, it is worth noting that some studies implement HRTF-independent testing for binaural localization models, as demonstrated by \textcite{zielinski_spatial_2022, zielinski_automatic_2022,antoniuk_blind_2023}.

@@

The optimizer was configured with the following hyperparameters: an initial learning rate of $10^{-3}$, a decay rate of $10^{-6}$, and momentum parameters $\beta_1 = 0.9$ and $\beta_2 = 0.999$. Training was conducted using a batch size of 8, with a maximum of 256 epochs set. An early stopping technique was implemented to prevent overfitting, terminating the process if no improvement was observed on the validation set for 20 consecutive epochs. Consequently, the maximum number of epochs was never reached; instead, training concluded after 25 to 36 epochs, with a median of 27.5 epochs. During the training process, the losses for both outputs were combined additively, ensuring equal weighting of both ensemble features.

The computational work for this study was conducted on a workstation equipped with an RTX Nvidia GeForce 4090 GPU and a 48-core AMD Ryzen ThreadRipper processor (up to 4.5 GHz). On the software side, MATLAB \parencite{MATLAB} with the Audio Toolbox \parencite{MATLAB_Audio_Toolbox} was used for the binaural recording synthesis, while Python \parencite{python} with the SciPy package \parencite{2020SciPy-NMeth} was used for feature extraction and Keras \parencite{chollet_keras_2015} for training the CNN model. The complete source code for all the experimental stages is publicly available on the GitHub repository \parencite{antoniuk_software_2024}. The spectrogram calculation phase required 21 minutes, data partitioning took 34 minutes, and the total training time for all iterations amounted to 40 minutes, making the entire training and evaluation process 95 minutes long.

@@

Figure \ref{fig:actual_vs_predicted_width} compares the actual and predicted ensemble widths for each sample, showing a heteroscedastic relationship between them, with a slight bias towards predicting lower ensemble width values for higher actual widths. This relationship exhibits a strong positive correlation, with a Pearson coefficient $r$ of 0.90. Additionally, the results indicate that the model provides more precise predictions for narrower ensemble widths, with an average MAE of $5.65^\circ$ for $\omega < 30^\circ$. However, performance deteriorates as the ensemble width increases, resulting in an MAE of $12.44^\circ$ for $\omega > 80^\circ$. This effect is more visible in Figure \ref{fig:mae_width}, which highlights the impact of the actual ensemble width on the precision of prediction. The correlation between the actual ensemble width and prediction error shows a weak positive relationship, with a Pearson coefficient $r$ of 0.27.

@@

The correlation between the actual and predicted ensemble location values exhibits a very high degree of correlation, as illustrated in Figure \ref{fig:actual_vs_predicted_location}. In this case, Pearson correlation coefficient $r$ is equal to as much as 0.97. In contrast to the ensemble width, no significant relationship is observed between actual location and its prediction error. This finding suggests that the model's ability to localize the center of the ensemble is robust, unaffected by the actual spatial positioning of the ensemble, including lateral locations. Figure \ref{fig:mae_location} corroborates this observation, demonstrating relatively consistent location error across most positions, with minor increases at extreme locations. The negligible correlation ($r = -0.03$) between the absolute location value and prediction error further supports the model's spatial invariance in its performance.

@@

Figure \ref{fig:map_mae_width} reveals a characteristic performance depression in $\omega \in [30\degree, 60\degree]$ previously shown from a different perspective in Figure \ref{fig:mae_width}. This heatmap highlights another interesting phenomenon in its upper corners as the error in these areas is considerably higher. This indicates that the model's performance for estimating ensemble width is substantially worse at extreme widths and locations, i.e., when both the width and locations are near their maximum investigated values ($|\theta| \approx 45\degree, \omega \approx 90\degree$).

@@

In terms of ensemble location prediction, the novelty of the proposed method makes direct comparison with existing literature challenging. However, its efficacy can be only evaluated indirectly against state-of-the-art individual-source localization techniques. The ensemble location prediction precision (MAE $=4.76\degree$) of the proposed method can be contextualized with the leading-edge binaural localization DeepEar model introduced by \textcite{yang_deepear_2022}. Their model reported MAEs of $7.4\degree$ and $2.3\degree$ for multi-source and single-source Angle of Arrival (AoA) estimation, respectively. As another promising example, the WaveLoc-CONV model developed by \textcite{vecchiotti_end--end_2019} demonstrated errors of $0\degree$ in anechoic conditions and $1.7\degree-2.4\degree$ in multi-condition scenarios. However, these results are limited to single-source speech localization, a substantially less complex task than the ensemble location prediction addressed by the proposed method. These experiments, while differing in objectives and datasets, provide valuable context for the proposed method's performance within current DoA and AoA estimation research.

@@

These limitations, however, present opportunities for future research. Despite the current constraints, this study introduces a novel method for characterizing acoustic scenes in binaural recordings of music, demonstrating substantial potential for advancing binaural audio analysis. The method offers promising prospects for developing innovative tools that can objectively analyze large repositories of binaural audio recordings, focusing on spatial content.

% \clearpage
\section*{Acknowledgments}
The work was supported by the grants from Białystok University of Technology (WI/WI-IIT/3/2022 and WZ/WI-IIT/5/2023) and funded with resources for research by the Ministry of Science and Higher Education in Poland.
